---
title: "Hobotnica_matrix_prep"
author: "Anton Zhelonkin"
date: "3/21/2022"
output: html_document
---

## Data preparation
Load data 
```{r}
control1 <- read.table(file = '/path/GSE163165_RAW/GSM4973752_1K_merged_gene_counts.txt',
                       sep = "\t", header = T)
lps1 <- read.table(file = '/path/GSE163165_RAW/GSM4973753_2K_merged_gene_counts.txt',
                   sep = "\t", header = T)
control2 <- read.table(file = '/path/GSE163165_RAW/GSM4973754_3K_merged_gene_counts.txt',
                       sep = "\t", header = T)
lps2 <- read.table(file = '/path/GSE163165_RAW/GSM4973755_4K_merged_gene_counts.txt',
                    sep = "\t", header = T)
control3 <- read.table(file = '/path//GSE163165_RAW/GSM4973756_5K_merged_gene_counts.txt',
                        sep = "\t", header = T)
lps3 <- read.table(file = '/path/GSE163165_RAW/GSM4973757_6K_merged_gene_counts.txt',
                    sep = "\t", header = T)
```

Checking the structure
```{r}
str(control1)
str(lps1)
```


Checking order
```{r}
all(control1 == lps1)
all(control2 == lps2)
all(control3 == lps3)
all(control1 == control2)
```
Length of columns

```{r}
length(control1$gene_name)
length(control2$gene_name)
length(control3$gene_name)
length(lps1$gene_name)
length(lps2$gene_name)
length(lps3$gene_name)
```

### Matrix preparation
Creating matrix  
                sample_1 ... sample_m  
gene_1  
gene_2  
....  
gene_n  
 
Merging columns onto matrix
```{r}
data_merge <- merge(control1, control2, by = c("Geneid", "gene_name") ) 
data_merge <- merge(data_merge, control3, by = c("Geneid", "gene_name") ) 
data_merge <- merge(data_merge, lps1, by = c("Geneid", "gene_name") ) 
data_merge <- merge(data_merge, lps2, by = c("Geneid", "gene_name") ) 
data_merge <- merge(data_merge, lps3, by = c("Geneid", "gene_name") ) 
head(data_merge)
```

ИNaming columns
```{r}
names(data_merge) <- c("gene_id", "gene_name", 
                       "control1", "control2",
                       "control3", "lps1",
                       "lps2", "lps3")
```

Checking structure
```{r}
str(data_merge)
```

## Cleaning dataframe to preapare analysis-ready matrix
```{r}
head(data_merge)
```
```{r}
expression_matrix <- data_merge[ , 2:8]
str(expression_matrix)
```
Deleting gene duplicates
```{r}
matrix_dupl_deleted = expression_matrix[!duplicated(expression_matrix$gene_name),]
matrix_dupl_deleted
length(expression_matrix$gene_name) - length(matrix_dupl_deleted$gene_name)
```

### Creating DE annotation of the experiment
Creating vectors
```{r}
sample_names_0 <- c("control1", "control2",
                    "control3", "lps1",
                    "lps2", "lps3")

condition_names_1 <- c("control", "control",
                    "control", "lps",
                    "lps", "lps")
```

Gatheting annotation
```{r}
annotation <- data.frame(sample_names_0, condition_names_1)
names(annotation) <- c(0, 1)
annotation
```

Сохраняю матрицу экспрессии
```{r}
write.table(file = "/path/matrix", x =  matrix_dupl_deleted, sep=",",  col.names=T)
```

Сохраняю аннотацию
```{r}
write.table(file = "/path/annotation", x =  annotation, sep=",",  col.names=T)
```



